---
title: "R Packages for Cointegration Testing with Structural Breaks, Fourier Approximations, and Causality Analysis"
abstract: >
  We present eight R packages implementing cointegration and Granger causality
  methods absent from the R ecosystem. The cointsmall package provides
  small-sample cointegration tests following Trinh (2022). The xtcadfcoint and
  xtbreakcoint packages implement panel cointegration tests with structural
  breaks and common factors. The hatemicoint and makicoint packages provide
  regime-shift cointegration tests with unknown break dates. The xtdhcoint
  package implements Durbin--Hausman panel cointegration tests. The fcoint
  package offers four Fourier-based cointegration tests. The caustests package
  provides a suite of Granger causality tests including Toda--Yamamoto, Fourier,
  quantile, and bootstrap variants. Together, these packages fill important gaps
  in the R cointegration toolkit, covering small-sample methods, structural
  breaks, cross-sectional dependence, Fourier approximations, and advanced
  causality testing. All packages are open-source under the GPL-3 license.
author:
  - name: Muhammad Abdullah Alkhalaf
    affiliation: Rufyq Elngeh for Academic and Business Services
    address: Riyadh, Saudi Arabia
    orcid: 0009-0002-2677-9246
    email: muhammedalkhalaf@gmail.com
type: package
output:
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: paper.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = "#>")
```

# Introduction

Cointegration testing is central to empirical macroeconomics and finance. The
classical Engle--Granger and Johansen tests, available in the `urca` package
\citep{Pfaff2008}, assume known break dates (or none), large samples, and
cross-sectional independence in panel settings. Several important advances
address these limitations:

- **Small-sample cointegration** tests valid with as few as 12 observations \citep{Trinh2022}.
- **Panel cointegration** with structural breaks and common factors \citep{Banerjee2015, Banerjee2024}.
- **Regime-shift cointegration** with unknown break dates \citep{HatemiJ2008, Maki2012}.
- **Durbin--Hausman panel tests** robust to cross-sectional dependence \citep{Westerlund2008}.
- **Fourier-based cointegration** tests capturing smooth structural change \citep{Banerjee2017fadl, Tsong2016}.
- **Advanced Granger causality** tests including Fourier, quantile, and bootstrap variants \citep{TodaYamamoto1995, Enders2016, Cai2023, Cheng2021}.

This article introduces eight R packages addressing these gaps.

# The cointsmall package

The `cointsmall` package implements the \citet{Trinh2022} cointegration test
designed for very small samples (as few as 12 observations) with up to two
endogenous structural breaks.

```{r cointsmall-example}
library(cointsmall)

# Simulate a short cointegrated series
set.seed(42)
n <- 30
x <- cumsum(rnorm(n))
y <- 0.8 * x + rnorm(n, 0, 0.3)

# Small-sample cointegration test with structural breaks
result <- cointsmall(y, x, breaks = 2, model = "cs",
                     criterion = "adf", trim = 0.15)
summary(result)
```

# The xtcadfcoint package

The `xtcadfcoint` package implements the \citet{Banerjee2024} panel cointegration
test with structural instabilities using cross-sectionally augmented Dickey--Fuller
regressions and Common Correlated Effects estimation.

```{r xtcadfcoint-example}
library(xtcadfcoint)

# Simulate panel data
set.seed(123)
N <- 15; T_obs <- 50
panel_data <- data.frame(
  country = rep(1:N, each = T_obs),
  year = rep(1:T_obs, N),
  y = unlist(lapply(1:N, function(i) {
    x <- cumsum(rnorm(T_obs))
    0.5 * x + rnorm(T_obs, 0, 0.3)
  })),
  x1 = unlist(lapply(1:N, function(i) cumsum(rnorm(T_obs)))),
  x2 = unlist(lapply(1:N, function(i) cumsum(rnorm(T_obs))))
)

result <- xtcadfcoint(y ~ x1 + x2, data = panel_data,
                      id = "country", time = "year",
                      model = 1, breaks = 1)
summary(result)
```

# The xtbreakcoint package

The `xtbreakcoint` package implements panel cointegration with structural breaks
and cross-section dependence following \citet{Banerjee2015}, with iterative
factor-break estimation and the \citet{BaiNg2004} MQ test.

```{r xtbreakcoint-example}
library(xtbreakcoint)

result <- xtbreakcoint(y ~ x1 + x2, data = panel_data,
                       id = "country", time = "year",
                       model = "trendshift", max_factors = 5)
summary(result)
```

# The hatemicoint package

The `hatemicoint` package implements the \citet{HatemiJ2008} cointegration test
with two unknown regime shifts, providing ADF*, Z$_t$*, and Z$_\alpha$*
statistics with endogenous break dates.

```{r hatemicoint-example}
library(hatemicoint)

set.seed(456)
n <- 150
x <- cumsum(rnorm(n))
y <- 0.5 * x + rnorm(n, 0, 0.4)

result <- hatemicoint(y, x, maxlags = 8,
                      lag_selection = "aic", trimming = 0.15)
summary(result)
```

# The xtdhcoint package

The `xtdhcoint` package implements the \citet{Westerlund2008} Durbin--Hausman
panel cointegration tests, robust to cross-sectional dependence via principal
components. It provides group-mean (DH$_g$) and panel (DH$_p$) statistics.

```{r xtdhcoint-example}
library(xtdhcoint)

result <- xtdhcoint(y ~ x1 + x2, data = panel_data,
                    id = "country", time = "year",
                    criterion = "bic")
summary(result)
```

# The fcoint package

The `fcoint` package provides four Fourier-based cointegration tests: FADL
\citep{Banerjee2017fadl}, FEG, FEG2, and the \citet{Tsong2016} KPSS-type test.

```{r fcoint-example}
library(fcoint)

set.seed(789)
n <- 200
x <- cumsum(rnorm(n))
y <- 1.2 * x + rnorm(n, 0, 0.5)

result <- fcoint(y, x, test = "fadl", max_freq = 3,
                 criterion = "aic")
print(result)
```

# The makicoint package

The `makicoint` package implements the \citet{Maki2012} test allowing an unknown
number of breaks (up to five) under four model specifications.

```{r makicoint-example}
library(makicoint)

result <- makicoint(y, x, model = 2, max_breaks = 3,
                    trimming = 0.10)
summary(result)
```

# The caustests package

The `caustests` package provides a comprehensive Granger causality suite:
Toda--Yamamoto \citep{TodaYamamoto1995}, Fourier single-frequency
\citep{Enders2016} and cumulative-frequency, quantile causality
\citep{Cai2023}, and bootstrap Fourier Granger causality in quantiles
\citep{Cheng2021}. The `test` argument (1--7) selects the method.

```{r caustests-example}
library(caustests)

set.seed(101)
n <- 200
x <- cumsum(rnorm(n))
y <- 0.3 * c(0, x[-n]) + rnorm(n)

# Fourier Toda-Yamamoto (test = 2)
result1 <- caustests(data = cbind(y, x), test = 2,
                     pmax = 4, kmax = 3, nboot = 1000)
summary(result1)

# Quantile causality (test = 6)
result2 <- caustests(data = cbind(y, x), test = 6,
                     quantiles = c(0.25, 0.5, 0.75), nboot = 500)
summary(result2)
```

# Comparison with existing implementations

The `urca` package \citep{Pfaff2008} provides Engle--Granger and Johansen
cointegration tests but does not support structural breaks, small samples, or
Fourier approximations. The `cointReg` package offers modified OLS estimators for
cointegrating regressions but not the specific test procedures implemented here.
The `egcm` package provides Engle--Granger cointegration with some model selection
but lacks break detection.

Our packages differ from existing implementations in several key ways:

- **cointsmall** is the only R package designed for cointegration testing with
  very small samples (n < 30).
- **xtcadfcoint** and **xtbreakcoint** provide panel cointegration tests with
  both structural breaks and cross-sectional dependence --- a combination
  unavailable in any existing R package.
- **hatemicoint** and **makicoint** implement regime-shift cointegration with
  endogenous break detection, methods available in Stata and GAUSS but not
  previously in R.
- **xtdhcoint** provides the Durbin--Hausman approach, which tests for
  cointegration rather than no-cointegration, offering a complementary null
  hypothesis to standard tests.
- **fcoint** captures smooth structural change via Fourier terms without
  requiring break date specification.
- **caustests** consolidates seven causality tests in a single interface,
  replacing the need for multiple packages or custom code.

# Summary

The eight packages presented here substantially extend R's cointegration and
causality testing capabilities. They address gaps in small-sample inference,
structural break detection, cross-sectional dependence, Fourier approximations,
and advanced causality testing. All packages are available on CRAN under GPL-3.

\bibliography{paper}
